---
output: 
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library('pds3')
```
# pds3

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/pds3)](https://cran.r-project.org/package=pds3)
[![Travis-CI Build Status](https://travis-ci.org/mwaldstein/pds3.svg?branch=master)](https://travis-ci.org/mwaldstein/pds3)
[![codecov.io](https://codecov.io/github/mwaldstein/pds3/coverage.svg?branch=master)](https://codecov.io/github/mwaldstein/pds3?branch=master)

pds3 reads [PDS3](https://pds.jpl.nasa.gov/datastandards/pds3/) files, a standard
published by JPL and used throughout NASA space missions. While PDS3 is being
supplanted by PDS4, a XML based standard, PDS3 is still being used and is
needed for accessing historic data.

## References

 * [PDS3 Data Standards](https://pds.jpl.nasa.gov/datastandards/pds3/)
 * [PDS3 Standards Reference](https://pds.jpl.nasa.gov/datastandards/pds3/standards/)
 * [PDS Data Archive](https://pds.jpl.nasa.gov/)

## Installation

pds3 is not yet stable enough to be on [CRAN](https://CRAN.R-project.org). You
can install the development version using devtools with:

``` r
devtools::install_github("mwaldstein/pds3")
```

## Basic Use

```{r example}
pds_label <- '
PDS_VERSION_ID                = PDS3

/* FILE CHARACTERISTICS */

RECORD_TYPE                   = FIXED_LENGTH
RECORD_BYTES                  = 3296
FILE_RECORDS                  = 3072

/* POINTERS TO DATA OBJECTS */

^IMAGE                        = "JNCR_2016345_03C00002_V01.IMG"

/* IDENTIFICATION DATA ELEMENTS */

FILE_NAME                     = "JNCR_2016345_03C00002_V01.IMG"
SPACECRAFT_NAME               = JUNO
MISSION_PHASE_NAME            = "SCIENCE ORBITS"
TARGET_NAME                   = JUPITER
INSTRUMENT_ID                 = "JNC"
INSTRUMENT_HOST_NAME          = "JUNO"
INSTRUMENT_NAME               = "JUNO EPO CAMERA"
PRODUCER_ID                   = JUNO_JUNOCAM_TEAM
DATA_SET_ID                   = "JUNO-J-JUNOCAM-3-RDR-L1A-V1.0"
STANDARD_DATA_PRODUCT_ID      = "JUNOCAM-RDR"
PRODUCT_CREATION_TIME         = 2017-05-31T18:42:49
PRODUCT_VERSION_ID            = "01"
PRODUCT_ID                    = "JNCR_2016345_03C00002_V01"
SOURCE_PRODUCT_ID             = "3D-0900010002-2016-349T17.19.13"
START_TIME                    = 2016-12-10T17:15:14.358
IMAGE_TIME                    = 2016-12-10T17:15:14.358
STOP_TIME                     = 2016-12-10T17:15:17.350
SPACECRAFT_CLOCK_START_COUNT  = "534662301:164"
SPACECRAFT_CLOCK_STOP_COUNT   = "N/A"

OBJECT                        = IMAGE
  LINES                       = 3072
  LINE_SAMPLES                = 1648
  SAMPLE_TYPE                 = UNSIGNED_INTEGER
  LINE_PREFIX_BYTES           = 0
  LINE_SUFFIX_BYTES           = 0
  SAMPLE_BITS                 = 16
  SAMPLE_BIT_MASK             = 2#1111111111111111#
  MD5_CHECKSUM                = "9ab98a1df127b82047e073e569cff24f"
END_OBJECT                    = IMAGE
END
'

lbl_object <- pds3_read(pds_label)
str(lbl_object)
```

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By participating in this project you agree to abide by its terms.
