---
title: "Getting Started With PDS3"
author: "Micah J Waldstein (micah@waldste.in)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started With PDS3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

PDS3 is a data standard used extensively by NASA for archiving data from science
missions, maintained by JPL. While being replaced by PDS4, PDS4 covers all
currently active missions and those covering the history of US space
exploration.

The R pds3 package provides tools for parsing PDS3 data, particularly the ODL
label format which describes all the metadata of data collection. Want to plot
a heatmap of Mars of all the images taken? This is the package for you!

## Getting Started

The package is not yet on CRAN, so to install you'll want to use `devtools` to
get the development version from Github.

```{r eval=FALSE}
# Install devtools if you don't already have it
install.packages("devtools")
# Install the development version of pds3
devtools::install_github("mwaldstein/pds3")
```

### Finding Data

If you're exploring this package, chances are you already have a collection of
PDS files, but if you're interested in exploring NASA data, the place to get
started is [The Planetary Data System](https://pds.jpl.nasa.gov/datasearch/data-search/).

Once you find a mission you're interested in exploring, finding the data
explorer is typically straight forward. The metadata this package processes is
typically stored in .lbl files. For instance, we'll look at the labels from the
Mars Reconnaissance Orbiter's HiRise experiment, particularly the data
associated with [this
image](http://ode.rsl.wustl.edu/mars/indexproductpage.aspx?product_idgeo=13465046).


## Reading & Parsing Data

We'll start by grabbing the label (LBL) file.

```{r fetch, cache=T}
href <- 'http://hirise.lpl.arizona.edu/PDS/RDR/ESP/ORB_011700_011799/ESP_011707_1440/ESP_011707_1440_COLOR.LBL'
req <- curl::curl_fetch_memory(href)
dat <- rawToChar(req$content)
```

Now we have the data, we'll process it using `parse_pds3`.

```{r process}
library(pds3)
res <- pds3_read(dat)
str(res$odl)
```

## Cautionary Notes

All of the data you can access from NASA missions is public, but be aware that
if you are going to download a large amount, there are almost always better
ways than web scraping - be it either a catalog-only download or via FTP,
investigate and be sure to be considerate.

Also of note that the data is public and free to use, it is the product of a
lot of hard work and it is appropriate to cite the data source and the
principle investigators of the instrument whose data you are using. NASA
provides a guide to [Citing PDS3
Data](https://pds.nasa.gov/datastandards/pds3/citing-pds3-data.shtml)

One cool thing is that many PDS catalogs include citation information in a PDS3
formatted file we can use this package to extract!

```{r fetch_catalog, cache=T}
cit_href <- "https://hirise.lpl.arizona.edu/PDS/CATALOG/RDR_DS.CAT"
cit_req <- curl::curl_fetch_memory(cit_href)
cit_dat <- rawToChar(cit_req$content)
cit_res <- pds3_read(cit_dat)

cit_res$odl$DATA_SET$DATA_SET_INFORMATION$CITATION_DESC
```

## References

 * The MRO-M-HIRISE-3-RDR-V1.0 data set was obtained from the [Planetary Data System (PDS)](https://pds.nasa.gov).
 * `r cit_res$odl$DATA_SET$DATA_SET_INFORMATION$CITATION_DESC`

